Chapter 7: Exercise 10
========================================================

### a
```{r 10a}
set.seed(1)
library(ISLR)
library(leaps)
attach(College)
train = sample(length(Outstate), length(Outstate)/2)
test = -train
College.train = College[train, ]
College.test = College[test, ]
reg.fit = regsubsets(Outstate~., data=College.train, nvmax=17, method="forward")
reg.summary = summary(reg.fit)
par(mfrow=c(1, 3))
plot(reg.summary$cp,xlab="Number of Variables",ylab="Cp",type='l')
min.cp = min(reg.summary$cp)
std.cp = sd(reg.summary$cp)
abline(h=min.cp+0.2*std.cp, col="red", lty=2)
abline(h=min.cp-0.2*std.cp, col="red", lty=2)
plot(reg.summary$bic,xlab="Number of Variables",ylab="BIC",type='l')
min.bic = min(reg.summary$bic)
std.bic = sd(reg.summary$bic)
abline(h=min.bic+0.2*std.bic, col="red", lty=2)
abline(h=min.bic-0.2*std.bic, col="red", lty=2)
plot(reg.summary$adjr2,xlab="Number of Variables",ylab="Adjusted R2",type='l', ylim=c(0.4, 0.84))
max.adjr2 = max(reg.summary$adjr2)
std.adjr2 = sd(reg.summary$adjr2)
abline(h=max.adjr2+0.2*std.adjr2, col="red", lty=2)
abline(h=max.adjr2-0.2*std.adjr2, col="red", lty=2)
```
All cp, BIC and adjr2 scores show that size 6 is the minimum size for the subset for which the scores are withing 0.2 standard deviations of optimum. We pick 6 as the best subset size and find best 6 variables using entire data.
```{r}
reg.fit = regsubsets(Outstate~., data=College, method="forward")
coefi = coef(reg.fit, id=6)
names(coefi)
```

### b
```{r 10b}
library(gam)
gam.fit = gam(Outstate ~ Private + s(Room.Board, df=2) + s(PhD, df=2) + s(perc.alumni, df=2) + s(Expend, df=5) + s(Grad.Rate, df=2), data=College.train)
par(mfrow=c(2, 3))
plot(gam.fit, se=T, col="blue")
```

### c
```{r}
gam.pred = predict(gam.fit, College.test)
gam.err = mean((College.test$Outstate - gam.pred)^2)
gam.err
gam.tss = mean((College.test$Outstate - mean(College.test$Outstate))^2)
test.rss = 1 - gam.err / gam.tss
test.rss
```
We obtain a test R-squared of 0.77 using GAM with 6 predictors. This is a slight improvement over a test RSS of 0.74 obtained using OLS. 

### d
```{r}
summary(gam.fit)
```
Non-parametric Anova test shows a strong evidence of non-linear relationship between response and Expend, and a moderately strong non-linear relationship (using p value of 0.05) between response and Grad.Rate or PhD. 