10
========================================================

### a

```{r 10a}
library(ISLR)
summary(Weekly)
pairs(Weekly)
cor(Weekly[,-9])
```

Year and Volume appear to have a relationship. No other patterns are
discernible.

### b

```{r}
attach(Weekly)
glm.fit = glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,
              data=Weekly,
              family=binomial)
summary(glm.fit)
```
Lag 2 appears to have some statistical significance with a Pr(>|z|) = 3%.

### c

```{r}
glm.probs = predict(glm.fit, type="response")
glm.pred = rep("Down", length(glm.probs))
glm.pred[glm.probs>.5] = "Up"
table(glm.pred, Direction)
```
Percentage of currect predictions: (54+557)/(54+557+48+430) = 56.1%. Weeks the 
market goes up the logistic regression is right most of the time, 557/(557+48)
= 92.1%. Weeks the market goes up the logistic regression is wrong most of the
time 54/(430+54) = 11.2%.

### d
```{r}
train = (Year < 2009)
Weekly.0910 = Weekly[!train,]
glm.fit = glm(Direction~Lag2,
              data=Weekly,
              family=binomial,
              subset=train)
glm.probs = predict(glm.fit, Weekly.0910,type="response")
glm.pred = rep("Down", length(glm.probs))
glm.pred[glm.probs>.5] = "Up"
Direction.0910 = Direction[!train]
table(glm.pred, Direction.0910)
mean(glm.pred == Direction.0910)
```

### e
```{r}
library(MASS)
lda.fit = lda(Direction ~ Lag2, data=Weekly, subset=train)
lda.pred = predict(lda.fit, Weekly.0910)
table(lda.pred$class, Direction.0910)
mean(lda.pred$class == Direction.0910)
```

## f
```{r}
qda.fit = qda(Direction~Lag2, data=Weekly, subset=train)
qda.class = predict(qda.fit, Weekly.0910)$class
table(qda.class, Direction.0910)
mean(qda.class == Direction.0910)
```
A correctness of 58.7% even though it picked Up the whole time!

### g
```{r}
library(class)
train.X = as.matrix(Lag2[train])
test.X = as.matrix(Lag2[!train])
train.Direction = Direction[train]
set.seed(1)
knn.pred = knn(train.X, test.X, train.Direction, k=1)
table(knn.pred, Direction.0910)
mean(knn.pred == Direction.0910)
```

### h
Logistic regression and LDA methods provide similar test error rates.

### i
```{r 10i}
# Logistic regression with Lag2:Lag1
glm.fit = glm(Direction~Lag2:Lag1, data=Weekly, family=binomial, subset=train)
glm.probs = predict(glm.fit, Weekly.0910, type="response")
glm.pred = rep("Down", length(glm.probs))
glm.pred[glm.probs>.5] = "Up"
Direction.0910 = Direction[!train]
table(glm.pred, Direction.0910)
mean(glm.pred == Direction.0910)

# LDA with Lag2 interaction with Lag1
lda.fit = lda(Direction ~ Lag2:Lag1, data=Weekly, subset=train)
lda.pred = predict(lda.fit, Weekly.0910)
mean(lda.pred$class == Direction.0910)

# QDA with sqrt(abs(Lag2))
qda.fit = qda(Direction~Lag2+sqrt(abs(Lag2)), data=Weekly, subset=train)
qda.class = predict(qda.fit, Weekly.0910)$class
table(qda.class, Direction.0910)
mean(qda.class == Direction.0910)

# KNN k =10
knn.pred = knn(train.X, test.X, train.Direction, k=10)
table(knn.pred, Direction.0910)
mean(knn.pred == Direction.0910)

# KNN k = 100
knn.pred = knn(train.X, test.X, train.Direction, k=100)
table(knn.pred, Direction.0910)
mean(knn.pred == Direction.0910)
```
Out of these permutations, the original LDA and logistic regression have better
performance in terms of test error rate.


