Chapter 6: Exercise 9
========================================================

### a
Load and split the College data
```{r}
library(ISLR)
set.seed(11)
sum(is.na(College))
train.size = dim(College)[1] / 2
train = sample(1:dim(College)[1], train.size)
test = -train
College.train = College[train, ]
College.test = College[test, ]
```

### b
NUmber of applications is the Apps variable.
```{r 9b}
lm.fit = lm(Apps~., data=College.train)
lm.pred = predict(lm.fit, College.test)
mean((College.test[, "Apps"] - lm.pred)^2)
```
Test RSS is $1538442$

### c
Pick $\lambda$ using College.train and report error on College.test
```{r 9c}
library(glmnet)
train.mat = model.matrix(Apps~., data=College.train)
test.mat = model.matrix(Apps~., data=College.test)
grid = 10 ^ seq(4, -2, length=100)
mod.ridge = cv.glmnet(train.mat, College.train[, "Apps"], alpha=0, lambda=grid, thresh=1e-12)
lambda.best = mod.ridge$lambda.min
lambda.best
ridge.pred = predict(mod.ridge, newx=test.mat, s=lambda.best)
mean((College.test[, "Apps"] - ridge.pred)^2)
```
Test RSS is slightly higher that OLS, $1608859$.

### d
Pick $\lambda$ using College.train and report error on College.test
```{r 9d}
mod.lasso = cv.glmnet(train.mat, College.train[, "Apps"], alpha=1, lambda=grid, thresh=1e-12)
lambda.best = mod.lasso$lambda.min
lambda.best
lasso.pred = predict(mod.lasso, newx=test.mat, s=lambda.best)
mean((College.test[, "Apps"] - lasso.pred)^2)
```
Again, Test RSS is slightly higher that OLS, $1635280$.

The coefficients look like
```{r}
mod.lasso = glmnet(model.matrix(Apps~., data=College), College[, "Apps"], alpha=1)
predict(mod.lasso, s=lambda.best, type="coefficients")
```

### e
Use validation to fit pcr
```{r 9e}
library(pls)
pcr.fit = pcr(Apps~., data=College.train, scale=T, validation="CV")
validationplot(pcr.fit, val.type="MSEP")
pcr.pred = predict(pcr.fit, College.train, ncomp=15)
mean((College.test[, "Apps"] - data.frame(pcr.pred))^2)
```

### f
Use validation to fit pls
```{r 9f}
pls.fit = plsr(Apps~., data=College.train, scale=T, validation="CV")
validationplot(pls.fit, val.type="MSEP")
pls.pred = predict(pls.fit, College.train, ncomp=10)
mean((College.test[, "Apps"] - data.frame(pcr.pred))^2)
```
