Chapter 8: Exercise 8
========================================================

### a
```{r 8a}
library(ISLR)
attach(Carseats)
set.seed(1)

train = sample(dim(Carseats)[1], dim(Carseats)[1] / 2)
Carseats.train = Carseats[train, ]
Carseats.test = Carseats[-train, ]
```

### b
```{r b8}
library(tree)
tree.carseats = tree(Sales~., data=Carseats.train)
summary(tree.carseats)
plot(tree.carseats)
text(tree.carseats, pretty=0)
pred.carseats = predict(tree.carseats, Carseats.test)
mean((Carseats.test$Sales - pred.carseats)^2)
```
The test MSE is about $4.15$.

### c
```{r 8c}
cv.carseats = cv.tree(tree.carseats, FUN=prune.tree)
par(mfrow=c(1, 2))
plot(cv.carseats$size, cv.carseats$dev, type="b")
plot(cv.carseats$k, cv.carseats$dev, type="b")

# Best size = 9
pruned.carseats = prune.tree(tree.carseats, best=9)
par(mfrow=c(1, 1))
plot(pruned.carseats)
text(pruned.carseats, pretty=0)
pred.pruned = predict(pruned.carseats, Carseats.test)
mean((Carseats.test$Sales - pred.pruned)^2)
```
Pruning the tree in this case increases the test MSE to $4.99$.

### d
```{r 9d}
library(randomForest)
bag.carseats = randomForest(Sales~., data=Carseats.train, mtry=10, ntree=500, importance=T)
bag.pred = predict(bag.carseats, Carseats.test)
mean((Carseats.test$Sales - bag.pred)^2)
importance(bag.carseats)
```
Bagging improves the test MSE to $2.58$. We also see that $\tt{Price}$, $\tt{ShelveLoc}$ and $\tt{Age}$ are three most important predictors of $\tt{Sale}$.

### e
```{r 9e}
rf.carseats = randomForest(Sales~., data=Carseats.train, mtry=5, ntree=500, importance=T)
rf.pred = predict(rf.carseats, Carseats.test)
mean((Carseats.test$Sales - rf.pred)^2)
importance(rf.carseats)
```
In this case, random forest worsens the MSE on test set to $2.87$. Changing $m$ varies test MSE between $2.6$ to $3$. We again see that $\tt{Price}$, $\tt{ShelveLoc}$ and $\tt{Age}$ are three most important predictors of $\tt{Sale}$.